Extended Data

### 2021-10-17 doc/manuscript/Dow_manuscript_Nature.Rmd
minor results edits

After fitting the model, we removed tree-years with poor fits. 
Models were judged to be poorly fit if certain modeled growth characteristics fell outside of the logical range. 
Modeled fits for tree-years were removed under five conditions:
(1) single day growth rates were higher than 2 standard deviations away from the mean for each wood-type (SCBI = 2, Harvard Forest = 34); 
(2) $DOY_{ip}$ was $\ge 2$ standard deviations away from the mean for it's xylem architecture group, year, and site (SCBI = 62, Harvard Forest = 106); 
(3) tree-years with small or negligible total growth ($\Delta DBH$ < 0.02 mm; SCBI = 0, Harvard Forest = 66); 
(4) model fit predicted total yearly growth to take longer than 365 days, indicating poor model fit (SCBI = 181, Harvard Forest = 199); 
and (5) models with unexplained sharp spikes in growth rate (SCBI = 0, Harvard Forest = 3). 
At Harvard Forest the tag years removed through this method were proportional to the original sample size, indicating that no species or size class was disproportionately removed compared to others. 
At SCBI a higher proportion of RP trees were removed--the majority falling under condition 4.

### Climate {-#Climate} 
For the dendrometer band analysis we used climate data from local weather stations. 
Daily maximum ($T_{max}$) and minimum temperatures ($T_{min}$) for SCBI were obtained from a meteorological tower adjacent to the SCBI ForestGEO plot, via the ForestGEO Climate Data Portal v1.0 [https://forestgeo.github.io/Climate/; @anderson-teixeira_forestgeo_2020]. 
Daily temperature records for Harvard Forest were obtained from the Harvard Forest weather station (Boose & Gould, 2004; Boose, 2018).
The R package climpact (no publication?) was used to plot temperatures for visual inspection. 
Climpact also identifies readings that were greater than 3 standard deviations away from yearly means. These points were labeled as outliers and removed from the dataset. 
Gaps in the SCBI meteorological tower data were subsequently filled using temperature readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail).

@@

Leaf phenology data for the years 2001-2018 were extracted for SCBI and Harvard Forest from the MCD12Q2 V6 Land Cover Dynamics product (informally called the MODIS Global Vegetation Phenology product) [@friedl_mcd12q2_2019]. This dataset was used to determine timing of greenup, midgreenup, senescence, along with enhanced vegetation index values for each year.

@@

The tree-ring records from our focal sites were complemented with a much larger collection spanning `r n_cores_sites-2` deciduous and mixed forest sites in Eastern North America[@maxwell_sampling_2020;@dorangeville_drought_2018] (Strange et al. 2019 (QUAL); Au et al. 2020 (ACSA); LeBlanc et al Accepted (LITU) ; Maxwell et al. 2020 (QUAL) ; Matheus and Maxwell 2018 ; Schmidt 2015 (Thesis), Maxwell et al. (2020); Maxwell and Harley 2017; Wilde and Maxwell 2018; Maxwell et al. 2019; Maxwell et al. 2015 --- *FIND REFS*).
Again, species were limited to broadleaf deciduous species with clearly defined xylem porosity (i.e., excluding semi-ring porous).

>>>>

After fitting the model, we removed tree-years with poor fits. 
Models were judged to be poorly fit if certain modeled growth characteristics fell outside of the logical range. 
Modeled fits for tree-years were removed under five conditions:
(1) single day growth rates were $\ge 2$ 2 standard deviations away from the mean for each wood-type (SCBI = 2, Harvard Forest = 34); 
(2) $DOY_{ip}$ was $\ge 2$ standard deviations away from the mean for it's xylem architecture group, year, and site (SCBI = 62, Harvard Forest = 106); 
(3) tree-years with small or negligible total growth ($\Delta DBH$ $\le 0.02$ mm; SCBI = 0, Harvard Forest = 66); 
(4) model fit predicted total yearly growth to take longer than 365 days, indicating poor model fit (SCBI = 181, Harvard Forest = 199); 
and (5) models with unexplained sharp spikes in growth rate (SCBI = 0, Harvard Forest = 3). 
At Harvard Forest the tag years removed through this method were proportional to the original sample size, indicating that no species or size class was disproportionately removed compared to others. 
At SCBI a higher proportion of RP trees were removed--the majority falling under condition 4.

### Climate {-#Climate} 
For the dendrometer band analysis we used climate data from local weather stations. 
Daily maximum ($T_{max}$) and minimum temperatures ($T_{min}$) for SCBI were obtained from a meteorological tower adjacent to the SCBI ForestGEO plot, via the ForestGEO Climate Data Portal v1.0 [https://forestgeo.github.io/Climate/; @anderson-teixeira_forestgeo_2020]. 
Daily temperature records for Harvard Forest were obtained from the Harvard Forest weather station (Boose & Gould, 2004; Boose, 2018).
The R package climpact <!--(no publication?)--> was used to plot temperatures for visual inspection. 
Climpact also identifies readings that were greater than 3 standard deviations away from yearly means. These points were labeled as outliers and removed from the dataset. 
Gaps in the SCBI meteorological tower data were subsequently filled using temperature readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail).

@@

Leaf phenology data for the years 2001-2018 were extracted for SCBI and Harvard Forest from the MCD12Q2 V6 Land Cover Dynamics product (informally called the MODIS Global Vegetation Phenology product) [@friedl_mcd12q2_2019]. This dataset was used to determine timing of greenup, midgreenup, senescence, and enhanced vegetation index values for each year.

@@

The tree-ring records from our focal sites were complemented with a much larger collection spanning `r n_cores_sites-2` deciduous and mixed forest sites in Eastern North America[@maxwell_sampling_2020;@dorangeville_drought_2018] (Strange et al. 2019 (QUAL); Au et al. 2020 (ACSA); LeBlanc et al Accepted (LITU) ; Maxwell et al. 2020 (QUAL) ; Matheus and Maxwell 2018 ; Schmidt 2015 (Thesis), Maxwell et al. (2020); Maxwell and Harley 2017; Wilde and Maxwell 2018; Maxwell et al. 2019; Maxwell et al. 2015 --- *FIND REFS - sent by Justin with tree cores*).
Again, species were limited to broadleaf deciduous species with clearly defined xylem porosity (i.e., excluding semi-ring porous).
